---
title: "R-Coding Lab"
author: "Dr. Matteo Tanadini"
date: "R-bootcamp"
output: 
  pdf_document:
    number_sections: true
header-includes:
  - \usepackage[bottom]{footmisc} 
  - \usepackage{color}
---


<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
\tableofcontents



<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
<!-- Setting global options (hidden section) -->

```{r globalOpts, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(concordance = TRUE,
               fig.align = "center",
               out.width = "0.7\\linewidth", ## default is 1
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               fig.pos = "H",
               crop = hook_pdfcrop) ## changes mar defaults...
##
options(digits = 2)
```



<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
# Coding
## Piping ("%>%")

```{r loadMagrittR}
library(magrittr)
## ("ceci n'est pas un pipe")
```

<!-- Note: when working with the *{dplyr}* package, the *{magrittr}* package is loaded by default. -->

```{r}
round(mean(iris$Sepal.Length), digits = 2)
## OR
## with "pipe" operator
iris$Sepal.Length %>% mean %>% round(digits = 2)
```


\bigskip
Same example, but with "exposition pipe" operator.

```{r}
iris %$% ## note the two types of pipe here
  mean(Sepal.Length) %>% 
  round(digits = 2)
```


\bigskip
A more complex example.

```{r}
mean(scale(iris$Petal.Length[iris$Species == "setosa"],
     scale = TRUE, center = FALSE))
## OR
## with "pipe" and "exposition pipe" operator
iris %>% 
  subset(Species == "setosa") %$% 
  scale(Petal.Length, scale = TRUE, center = FALSE) %>% 
  mean
```


## Reshaping data sets

```{r}
# d.sport <- read.table("DataSport.txt", header = TRUE,
#                       nrows = 7)
# ## OR
d.sport <-read.table(
  "http://stat.ethz.ch/Teaching/Datasets/WBL/sport.dat",
  header = TRUE, nrows = 7)
##
d.sport
str(d.sport)
```


\bigskip
First take rownames as an actual variable ("there is no such a thing as 'metadata'").

```{r}
library(tibble) ## for rownames_to_column()
d.sport <- d.sport %>% 
  rownames_to_column(var = "Athlete")
d.sport
```


\bigskip
In this data set each row represents an athlete. Each athlete has a record for each discipline. This data is said to be in wide-format.

Let's turn this data set into a long-format data set. Each row will then represent the performance of an athlete in a given discipline (e.g. weit for OBRIEN).

```{r}
library(tidyr) ## for gather()
d.sport.long <- gather(d.sport, 
                 key = "discipline", ## new column with name of the discipline
                 value = "result", ## new column with value (unquoted also works)
                 -Athlete) ## variable(s) that is(are) not to put as results
##
head(d.sport.long)
d.sport.long %>% 
  subset(subset = Athlete == "OBRIEN")
## OR
# library(dplyr)
# d.long %>% 
#   filter(Athlete == "OBRIEN")
```


\bigskip
Let's get back to a wide data set.

```{r}
d.sport.wide.again <- spread(d.sport.long, 
                             key = "discipline", ## name from long df
                             value = "result") ## name from long df
head(d.sport.wide.again)
```


\bigskip

**Note:** the data format needed for supervised problems is long-format (each row represents a single observation). 
Whereas, for unsupervised problems, wide-format is required (i.e. one represents several measurements). For example, a sample of wine where several aspects where quantified.

<!-- Change from upper case names to lower case names. -->
<!-- ```{r} -->
<!-- d.sport.wide.again$Athlete <- tolower(d.sport.wide.again$Athlete) -->
<!-- head(d.sport.wide.again) -->
<!-- ``` -->


## Joining data sets


```{r}
set.seed(14)
d.age <- data.frame(age = runif(n = 7, min = 19, max = 34),
                    Athlete = d.sport$Athlete,
                    Gender = c("M","M","F","M","F","F","F"))
d.age
```


\bigskip
Let's add this information to the long-format data set.

```{r}
library(dplyr)
d.sport.long.age <- left_join(d.sport.long, d.age, by = "Athlete")
head(d.sport.long.age)
```

\bigskip

```{r}
class(d.sport.long$Athlete)
class(d.age$Athlete)
```


\bigskip
If names of the shared variable differs between the two data sets.

```{r}
set.seed(14)
d.age.2 <- data.frame(age = runif(n = 7, min = 19, max = 34),
                      person = d.sport$Athlete)
```

\bigskip

```{r}
d.sport.long.age.2 <- left_join(d.sport.long, d.age.2, 
                              by = c("Athlete" = "person"))
head(d.sport.long.age.2)
```


\bigskip
see *?join* for further function to join two data sets together.


## Warning about *{dplyr}* and loops

Avoid using *{dplyr}* functions within loops (e.g. *for* loops).








